---
title: "Top 10 Contributing Californian Cities in 2016 Election"
output: html_document
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, include=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(here)
library(skimr)
library(janitor)
library(httr)
library(readxl)
library(vroom)
library(infer)
```
```


We would like to produce the plot that shows the top ten cities in highest amounts raised in political contributions in California during the 2016 US Presidential election.

To get this plot, we must join two dataframes; the one you have with all contributions, and data that can translate zipcodes to cities. We can find a file with all US zipcodes, e.g., here http://www.uszipcodelist.com/download.html.

The easiest way would be to create two plots and then place one next to each other. For this, you will need the patchwork package. https://cran.r-project.org/web/packages/patchwork/index.html

Let's first looks at the data.

```{r load_data}
CA_contributors_2016 <- vroom::vroom(here::here("data","CA_contributors_2016.csv"))
glimpse(CA_contributors_2016)
```

To link zipcode to cities, we need another data source.

```{r load_zip}
City_zipcode <- vroom::vroom(here::here("data","zip_code_database.csv"))
glimpse(City_zipcode)
```

Then we need to join the two dataframes to pair cities and zipcodes, while ignore other unrelevant data.

```{r city_zip}
City_zip <- City_zipcode %>% 
  select(zip, primary_city)%>% #only zip and city are relevant in the task
  mutate(zip=as.double((zip))) #convert zip in City_zipcode into double as zip in CA_contributors_2016 is double

glimpse(City_zip)
```

Now, we plot the top contributing cities for the top 2 candidates in the election.
```{r top_2}
#join the two dataframes
contributors_cities<-full_join(CA_contributors_2016,City_zip)

#Hillary Clinton's top contributing cities
top10_HRC <- contributors_cities %>% 
  filter(cand_nm=="Clinton, Hillary Rodham") %>% 
  group_by(primary_city) %>% 
  summarize(amount=sum(contb_receipt_amt)) %>% #sum of amount contributed by each city
  arrange(desc(amount)) %>% 
  slice(1:10) #top 10 by amount

#Trump's top contributing cities
top10_DJT <- contributors_cities %>% 
  filter(cand_nm=="Trump, Donald J.") %>% 
  group_by(primary_city) %>% 
  summarize(amount=sum(contb_receipt_amt)) %>% #sum of amount contributed by each city 
  arrange(desc(amount)) %>% 
  slice(1:10) #top 10 by amount

#plot for Hilary Clinton
p1 <- ggplot(top10_HRC, aes(x=amount, y=reorder(primary_city,amount)))+ #arrange in descending order of amount
  geom_col(fill='#2E74C0') + #color the series in blue
  scale_x_continuous(breaks = seq(0, 12000000, by = 4000000),labels = label_comma(prefix = "$"))+ #formatting x-axis (max, min and accounting style)
  facet_wrap(~"Clinton, Hillary Rodham")+ #adding grey box to candidate name
  theme_bw() + #black outline and grey grids
  theme(axis.title.y = element_blank(),axis.title.x = element_blank(),plot.title = element_text(hjust=0.5,size=10))+ #remove axis labels and format title
NULL

p2 <- ggplot(top10_DJT, aes(x=amount, y=reorder(primary_city,amount))) +#arrange in descending order of amount 
  geom_col(fill='#CB454A') + #color the series in red
  scale_x_continuous(breaks = seq(0, 400000, by = 200000),labels = label_comma(prefix = "$") )+ #formatting x-axis (max, min and accounting style)
  facet_wrap(~"Trump, Donald J.")+ #adding grey box to candidate name
  theme_bw()+theme(axis.title.y = element_blank(),axis.title.x = element_blank(),plot.title = element_text(hjust=0.5,size=10))+ #remove axis labels and format title
NULL


patch=p1+p2 #put p1 and p2 together

#format joined plot - title on top centered and caption at the bottom centered
patch+plot_annotation(title="Where did candidates raise most money?", caption ="Amount raised",theme=theme(plot.title=element_text(hjust=0.25,size=13),plot.caption = element_text(hjust=0.5)))
```

Now, let's find out about the top 10 candidates and present them together.

```{r top_10}
A <- Top_10_cities %>% 
    filter(cand_nm=="Bush, Jeb")

B <- Top_10_cities %>% 
    filter(cand_nm=="Carson, Benjamin S.")

C <- Top_10_cities %>% 
    filter(cand_nm=="Cruz, Rafael Edward 'Ted'")

D <- Top_10_cities %>% 
    filter(cand_nm=="Fiorina, Carly")

E<- Top_10_cities %>% 
    filter(cand_nm=="Paul, Rand")

F <- Top_10_cities %>% 
    filter(cand_nm=="Rubio, Marco")

G <- Top_10_cities %>% 
    filter(cand_nm=="Sanders, Bernard")

p3<- ggplot(A, aes(x=raised, y=reorder(primary_city,raised)))+ #arrange in descending order of amount raised
  geom_col(fill='#C0C0C0')+
  facet_wrap(~"Bush, Jeb")+
  scale_x_continuous(labels = label_comma(prefix = "$"))+ #formatting x-axis (max, min and accounting style)
  theme_bw() + #black outline and grey grids
  theme(axis.title.y = element_blank(),axis.title.x = element_blank(),plot.title = element_text(hjust=0.5,size=10))+ #remove axis labels and format title
  NULL

p4<- ggplot(B, aes(x=raised, y=reorder(primary_city,raised)))+ #arrange in descending order of amount raised
  geom_col(fill='#808080')+
  facet_wrap(~"Carson, Benjamin S.")+
  scale_x_continuous(labels = label_comma(prefix = "$"))+ #formatting x-axis (max, min and accounting style)
  theme_bw() + #black outline and grey grids
  theme(axis.title.y = element_blank(),axis.title.x = element_blank(),plot.title = element_text(hjust=0.5,size=10))+ #remove axis labels and format title
  NULL

p5<- ggplot(C, aes(x=raised, y=reorder(primary_city,raised)))+ #arrange in descending order of amount raised
  geom_col(fill='#800000')+
  facet_wrap(~"Cruz, Rafael Edward 'Ted'")+
  scale_x_continuous(labels = label_comma(prefix = "$"))+ #formatting x-axis (max, min and accounting style)
  theme_bw() + #black outline and grey grids
  theme(axis.title.y = element_blank(),axis.title.x = element_blank(),plot.title = element_text(hjust=0.5,size=10))+ #remove axis labels and format title
  NULL

p6<- ggplot(D, aes(x=raised, y=reorder(primary_city,raised)))+ #arrange in descending order of amount raised
  geom_col(fill='#808000')+
  facet_wrap(~"Fiorina, Carly")+
  scale_x_continuous(labels = label_comma(prefix = "$"))+ #formatting x-axis (max, min and accounting style)
  theme_bw() + #black outline and grey grids
  theme(axis.title.y = element_blank(),axis.title.x = element_blank(),plot.title = element_text(hjust=0.5,size=10))+ #remove axis labels and format title
  NULL

p7<- ggplot(E, aes(x=raised, y=reorder(primary_city,raised)))+ #arrange in descending order of amount raised
  geom_col(fill='#008000')+
  facet_wrap(~"Paul, Rand")+
  scale_x_continuous(labels = label_comma(prefix = "$"))+ #formatting x-axis (max, min and accounting style)
  theme_bw() + #black outline and grey grids
  theme(axis.title.y = element_blank(),axis.title.x = element_blank(),plot.title = element_text(hjust=0.5,size=10))+ #remove axis labels and format title
  NULL

p8<- ggplot(F, aes(x=raised, y=reorder(primary_city,raised)))+ #arrange in descending order of amount raised
  geom_col(fill='#800080')+
  facet_wrap(~"Rubio, Marco")+
  scale_x_continuous(labels = label_comma(prefix = "$"))+ #formatting x-axis (max, min and accounting style)
  theme_bw() + #black outline and grey grids
  theme(axis.title.y = element_blank(),axis.title.x = element_blank(),plot.title = element_text(hjust=0.5,size=10))+ #remove axis labels and format title
  NULL

p9<- ggplot(G, aes(x=raised, y=reorder(primary_city,raised)))+ #arrange in descending order of amount raised
  geom_col(fill='#008080')+
  facet_wrap(~"Sanders, Bernard")+
  scale_x_continuous(labels = label_comma(prefix = "$"))+ #formatting x-axis (max, min and accounting style)
  theme_bw() + #black outline and grey grids
  theme(axis.title.y = element_blank(),axis.title.x = element_blank(),plot.title = element_text(hjust=0.5,size=10))+ #remove axis labels and format title
  NULL

patched=p1+p2+p3+p4+p5+p6+p7+p8+p9

patched+
plot_annotation(title="Where did candidates raise most money?", caption ="Amount raised",theme=theme(plot.title=element_text(hjust=0.05,size=13),plot.caption = element_text(hjust=0.5)))+theme(axis.title.y = element_blank(),axis.title.x = element_blank())
```

Voila!



